---
title: "conflictABM"
author: "Clara Lavrador"
date: "2026-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview

### Life Cycle

1. Pursue
2. Initial Attacks
3. Contingent Attacks
4. Game
5. Update Enmity 

#### Pursue

Agents will combine other agents' association value with their enemy likelihood to produce a score for each agent.

The n number with the highest scores will be the agent's "pursuits".

The n number with the lowest scores will be the agent's "enemies".

#### Initial Attacks

Agents will decide whether to attack each other agent.

The probability of attack will be higher for those that the agent defined as "enemies".

#### Contingent Attacks

Agents will be shown all initial attacks involving the agents they pursued. 

They will fill whatever triadic primitive role is available where they are grouped with their pursuit.

For those that their pursuits attacked, they will complete alliance or displacement.

For those that attacked their pursuits, they will complete defense or generalization.

Which of the two they complete depends on which is available. No double attacks (same attack from Intial Attacks round) are permitted. If one triad already exists from the Initial Attacks round, the other primitive will be completed. If both already exist, nothing happens. If neither exists, one is randomly selected.

#### Game


#### Update Enmity

Agents will update the likelihood that each other agent is their enemy based on the previous attacks. 



```{r packages}

# library(dplyr)
# library(igraph)
# library(tidyverse)
```


#Parameters
```{r parameters}

#number of agents
popSize = 50

#number of pursuits agents can make
numPursuits = 4

#number of enemies
numEnemies = 4

#number of rounds
rounds = 300

#probability of attacking enemy
probAttackEnemy = 0.7

#probability of attacking nonenemy
probAttackNon = 0.2

#number of generations
generations = 1

#number of loops
modelLoops = 1

multiplier = 1.4

#strength of selection
selStrength = 10

#probability of a strat flipping
flipProb = 0.002

#prim combo abbreviations 
#list the prim combo abbreviations in order to append later
primCombos <- c("A",
"P",
"D",
"G",
"AP",
"AD",
"AG",
"PD",
"PG",
"DG",
"APD",
"APG",
"ADG",
"PDG",
"APDG",
"XX")

primComboDF <- data.frame("combo" =  primCombos,
                             "tot" = NA)

primComboDF$tot <- nchar(primComboDF$combo)

primComboDF$tot[primComboDF$combo == "XX"] <- 0

#we need popSize/5 of each prim tot
tots <- unlist(lapply(unique(primComboDF$tot), function(i) {
  sample(primComboDF$combo[primComboDF$tot == i], 
         popSize/5, 
         replace = TRUE)
}))

#write the path the CSV will save to
#masterPath <- "G:/My Drive/groupFormationResults/Actual Results/masterResultsM.csv"

```

#Functions
```{r functions}

#Agent generate function#

agentGenerate <- function(n){

  #agent IDs
  id <- c(1:n)

  #association value
  av <- round(abs(rnorm(n, mean = 50, sd = 10)), digits = 2)
  av <- av/100

  #strats
  # alliance <- rbinom(n, 1, 0.5)
  # generalization <- rbinom(n, 1, 0.5)
  # displacement <- rbinom(n, 1, 0.5)
  # defense <- rbinom(n, 1, 0.5)
  
  alliance <- 1
  generalization <- 1
  displacement <- 1
  defense <- 1

  # alliance <- c(rep(c(0,1), each = popSize/2))
  # generalization <- c(rep(c(0,1), each = popSize/2))
  # displacement <- c(rep(c(0,1), each = popSize/2))
  # defense <- c(rep(c(0,1), each = popSize/2))

  payoff <- 0

  agents <- data.frame(id, av, alliance, generalization, displacement, defense, payoff)

  return(agents)

}


#diff agent generate function that makes number of prims in strat equal in population
# agentGenerate <- function(n){
# 
#   #agent IDs
#   id <- c(1:n)
# 
#   #association value
#   av <- round(abs(rnorm(n, mean = 50, sd = 10)), digits = 2)
#   av <- av/100
#   
#   #strategy
#   combo <- sample(tots, size = n, replace = F)
# 
#   #strats
#   alliance <- NA
#   generalization <- NA
#   displacement <- NA
#   defense <- NA
# 
#   payoff <- 0
# 
#   agentDF <- data.frame(id, av, combo, alliance, generalization, displacement, defense, payoff)
#   
#   #map each letter to its prim
#   primMap <- c(
#     alliance = "A",
#     generalization = "G",
#     displacement = "P",
#     defense = "D"
#   )
#   
#   #fill in binary prim cols based on letters in combo
#   agentDF[names(primMap)] <-
#     sapply(primMap, function(letter) {
#       as.integer(grepl(letter, agentDF$combo))
#     })
# 
#   #delete combo col
#   agentDF <- agentDF[-c(3)]
# 
#   return(agentDF)
# 
# }


#Association Matrix#
#make association mat so that there is noise in how AVs are perceived by others

makeAssociationMat <- function(agentsDF, sd){
  
  associationMat <- matrix(, nrow = popSize, ncol = popSize)
  
  for(a in 1:nrow(agentsDF)){
    
    associationMat[,a] <-  rnorm(popSize, mean = agentsDF$av[a], sd = sd)
    
  }
  
  diag(associationMat) <- NA
  
  return(associationMat)
  
}



#Transitive Inference#
#Optimized version of transUpdate

transUpdate <- function(agent, agentsDF, pursuitMatrix, attackMatrix, enemyMatrix){
  
  #make the attack matrix diagonal NA
  diag(attackMatrix) <- NA
  
  #get the prims they use
  primCols <- c("alliance", "generalization", "displacement", "defense")
  primStrat <- primCols[agentsDF[agent, primCols] == 1]
  
  #pull the indices of the agents that the focal agent hates most
  enemies <- which(pursuitMatrix[agent, ] == -1)
  
  #vector of enemies that focal agent attacked
  sentAttacks <- enemies[attackMatrix[agent,enemies] == 1]
  
  #vector of enemies that attacked focal agent 
  recAttacks <- enemies[attackMatrix[enemies,agent] == 1] 
  
  posUpdate <- c()
  negUpdate <- c()
  
  #Alliance#
  #focal agent asks: who else attacked the agents I attacked?
  if("alliance" %in% primStrat && length(sentAttacks) > 0){
    subMat <- attackMatrix[, sentAttacks, drop = FALSE]
    posUpdate <- c(posUpdate, which(subMat == 1, arr.ind = TRUE)[, "row"])
    negUpdate <- c(negUpdate, which(subMat == 0, arr.ind = TRUE)[, "row"])
  }
  
  #Displacement#
  #focal agent asks: who did the agents I attacked attack?
  if("displacement" %in% primStrat && length(sentAttacks) > 0){
    subMat <- attackMatrix[sentAttacks, , drop = FALSE]
    posUpdate <- c(posUpdate, which(subMat == 1, arr.ind = TRUE)[, "col"])
    negUpdate <- c(negUpdate, which(subMat == 0, arr.ind = TRUE)[, "col"])
  }
  
  #Defense#
  #focal agent asks: who attacked the agents who attacked me?
  if("defense" %in% primStrat && length(recAttacks) > 0){
    subMat <- attackMatrix[, recAttacks, drop = FALSE]
    posUpdate <- c(posUpdate, which(subMat == 1, arr.ind = TRUE)[, "row"])
    negUpdate <- c(negUpdate, which(subMat == 0, arr.ind = TRUE)[, "row"])
  }
  
  #Generalization#
  #focal agent asks: who else did the agents who attacked me attack?
  if("generalization" %in% primStrat && length(recAttacks) > 0){
    subMat <- attackMatrix[recAttacks, , drop = FALSE]
    posUpdate <- c(posUpdate, which(subMat == 1, arr.ind = TRUE)[, "col"])
    negUpdate <- c(negUpdate, which(subMat == 0, arr.ind = TRUE)[, "col"])
  }
  
  #remove self
  posUpdate <- posUpdate[posUpdate != agent]
  negUpdate <- negUpdate[negUpdate != agent]
  
  #update for each occurrence
  
  #neg update (increase enemy prob)
  if(length(negUpdate) > 0){
    for(agentName in negUpdate){
      prior <- enemyMatrix[agent, agentName]
      num <- prior * (1 - probAttackNon)
      enemyMatrix[agent, agentName] <- num / (num + (1 - prior) * (1 - probAttackEnemy))
    }
  }
  
  #pos update (decrease enemy prob)
  if(length(posUpdate) > 0){
    for(agentName in posUpdate){
      prior <- enemyMatrix[agent, agentName]
      num <- prior * probAttackNon 
      enemyMatrix[agent, agentName] <- num / (num + (1 - prior) * probAttackEnemy)
    }
  }
  
  return(enemyMatrix)
  
}


contingentAttack <- function(agent, attackMatrix, pursuitMatrix){
  
  #pull the agent's pursuits
  pursuits <- which(pursuitMatrix[agent,] == 1)
  
  #subset attack matrix to get just the attacks
  idx <- which(attackMatrix == 1, arr.ind = TRUE)
  
  #pull those who pursuits attacked
  victims <- idx[idx[, "row"] %in% pursuits, , drop = F]
  victims <- unique(victims[,"col"])
  
  #pull those who attacked pursuits
  attackers <- idx[idx[, "col"] %in% pursuits, , drop = F]
  attackers <- unique(attackers[, "row"])
  
  #remove self, in case agent was attacked by/attacked the agent they pursued
  victims <- victims[victims != agent]
  attackers <- attackers[attackers != agent]
  
  if (length(victims > 0)) {
    
    #loop through each victim
    for (i in 1:length(victims)) {
      
      #get the ID of the victim
      vic = victims[i]
      
      #if the agent already attacked that victim...
      if (attackMatrix[agent, vic] %in% c(1, 2) &
          attackMatrix[vic, agent] == 0) {
        
        #that victim will attack the agent
        attackMatrix[vic, agent] <- 2
        
        #if the agent was already attacked by that victim...
      } else if (attackMatrix[agent, vic] == 0 &
                 attackMatrix[vic, agent] %in% c(1, 2)) {
        
        #then agent will attack the victim
        attackMatrix[agent, vic] <- 2
        
        #if there were no attacks between them
      } else if (attackMatrix[agent, vic] == 0 &
                 attackMatrix[vic, agent] == 0) {
        
        #randomly choose which will happen
        ifelse(rbinom(1, 1, 0.5) == 0, attackMatrix[vic, agent] <- 2, attackMatrix[agent, vic] <- 2)
        
      }
      
    }
    
  }
  
  if (length(attackers > 0)) {
    
    #loop through each attacker
    for (i in 1:length(attackers)) {
      
      #get the ID of the attacker
      att = attackers[i]
      
      #if the agent already attacked that attacker...
      if (attackMatrix[agent, att] %in% c(1, 2) &
          attackMatrix[att, agent] == 0) {
        
        #that attacker will attack the agent
        attackMatrix[att, agent] <- 2
        
        #if the agent was already attacked by that attacker...
      } else if (attackMatrix[agent, att] == 0 &
                 attackMatrix[att, agent] %in% c(1, 2)) {
        
        #then agent will attack the attacker
        attackMatrix[agent, att] <- 2
        
        #if there were no attacks between them
      } else if (attackMatrix[agent, att] == 0 &
                 attackMatrix[att, agent] == 0) {
        
        #randomly choose which will happen
        ifelse(rbinom(1, 1, 0.5) == 0, attackMatrix[att, agent] <- 2, attackMatrix[agent, att] <- 2)
        
      }
      
    }
    
  }

  return(attackMatrix)
  
}

#note that here I'm using a 2 to indicate the round 2 attacks. that way, the two agents involved
#in the attack are able to see that that happened to prevent a double attack, while obscuring 
#these round 2 attacks from third parties (only subset attackMat == 1)
###might be interesting to keep the 2s and allow agents to distinguish contingent attacks,
#but for now just turning the 2s into 1s once everyone is done.



#Contributions#

payoff <- function(gameMatrix, gamers, agentDF){
  
  #start the sum
  contributionTot <- 0
  
  #for each agent...
  for(agent in gamers){
    
    #their max contribution is their association value
    maxContribution <- agentDF$av[row.names(agentDF) == agent] * 100
    
    #get the percent of the AV they will contribute, which is proportional to number of recip ties within group
    proportionTied <- if (nrow(gameMatrix) > 1) sum(gameMatrix[as.character(agent), ])/(nrow(gameMatrix)-1) else 0
    
    #calculate their contribution 
    contribution <- maxContribution * proportionTied
    
    #add their total contribution to those of the other agents
    contributionTot <- contributionTot + contribution
    
    #add their noncontributed pay to their corresponding slot in payoff vector (what they keep)
    agentDF$payoff[as.numeric(agent)] <- agentDF$payoff[as.numeric(agent)] + (maxContribution - contribution)
    
  }
  
  contributionTot <- contributionTot * multiplier
  
  #divide the pooled contributions by the number of agents in the group
  basePay <- contributionTot/nrow(gameMatrix)
  
  agentDF$payoff[as.numeric(gamers)] <- agentDF$payoff[as.numeric(gamers)] + basePay
  
  #return the payoff vector
  return(agentDF)
  
}

#Reproduction#

reproduce <- function(agentDF, selStrength, n){
  
  #rescale payoffs to be between 0-1
  agentDF$payoff <- agentDF$payoff - min(agentDF$payoff)
  agentDF$payoff <- agentDF$payoff/max(agentDF$payoff)
  
  #add constant to prob weight to fix strength of selection
  agentDF$payoff <- agentDF$payoff + (1/selStrength) #the bigger the decimal, the larger the diff in probability
  
  #generate random offspring
  offspringDF <- agentDF[sample(1:nrow(agentDF), n, replace = T, prob = agentDF$payoff),] #the bigger the decimal, the larger the diff
  
  #Reset IDs
  row.names(offspringDF) <- 1:n
  
  offspringDF$id <- 1:n
  
  #Reset payoff vec
  offspringDF$payoff <- 0
  
  #Mutation:
  
  #for each agent...
  for (a in 1:nrow(offspringDF)) {
    
    #change their association value to be within certain SD of original parent AV
    association_value <- rnorm(1, mean = offspringDF$av[a], sd = .01)
    
    #don't let association value get above 1
    association_value <- ifelse(association_value > 1, 1 - (association_value - 1), association_value)
    
    #don't let association value get below 0
    association_value <- ifelse(association_value < 0, 0 + abs(association_value), association_value)
    
    #assign the new mutated association value
    offspringDF$av[a] <- association_value
    
    #flip each trait or nah and replace into corresponding cells
    offspringDF[a, 3:6] <- (offspringDF[a, 3:6] + rbinom(4, 1, flipProb)) %% 2
    
   }
  
  offspringDF <- offspringDF[-c(8)]
  
  return(offspringDF)
  
}






```

#Life Cycle
```{r lifeCycle} 

#analysis dataframe
# results <- data.frame(
#   "loop" = rep(1:modelLoops, each = generations*length(primCombos)),
#   "generation" = rep(1:generations),
#   "primCombo" = rep(primCombos))


resultsList <- list()
counter <- 1

# START #

for (m in 1:modelLoops) {
  
  print(m)
  
  #birth the agents
  agents <- agentGenerate(popSize)
  
  # GENERATION #
  
  for (g in 1:generations) {
    
    print(g)
    
    #generate association mat so agents see each other's av a bit differently
    associationMat <- makeAssociationMat(agents, 0.10)
    
    #matrix that holds each agent's perceived enemy status with each other agent
    enemyMat <- matrix((numEnemies / popSize), nrow = popSize, ncol = popSize)
    diag(enemyMat) <- NA
    
    #this is just so i can see what's going on; not referenced
    enemyMatList <- list()
  
    # #pursuit matrix list of noncumulative interactions
    #  pursuitMat <- matrix(0, nrow = popSize, ncol = popSize)
    #  diag(pursuitMat) <- NA
    pursuitMatList <- list()
    # 
    # #matrix that holds attacks to be backreferenced for previous rounds
    # attackMatPrev <- matrix(0, nrow = popSize, ncol = popSize)
    # diag(attackMatPrev) <- NA

    
    # ROUNDS #
    
    for (r in 1:rounds) {
      
      #make the empty new pursuit matrix
      pursuitMat <- matrix(0, nrow = popSize, ncol = popSize)
      diag(pursuitMat) <- NA
      
      #each agent chooses k other agents to pursue
      for (a in 1:popSize) {
        
        if(r == 1){
          
          groupVals <- 1
          
        } else {
          
          groupVals <- 1.01 - enemyMat[a, ]
          
        }
        
        #pull others' association values from the perspective of a
        assocVals <- associationMat[a, ]
        
        #calculate the pursuit scores by combining association and reciprocity
        pursuitScores <- assocVals * groupVals
        
        #nullify their own
        pursuitScores[a] <- NA
        
        #order from highest to lowest scores
        pursuit_indices <- order(-pursuitScores)[1:numPursuits]
        
        #add 1 to their pursuits in the pursuit matrix
        pursuitMat[a, pursuit_indices] <- 1
        
        #order from lowest to highest scores
        enemies_indices <- order(pursuitScores)[1:numEnemies]
        
        #add -1 to their enemies in the pursuit matrix
        pursuitMat[a, enemies_indices] <- -1
        
      }
      
      #bind the pursuit mat to the list
      pursuitMatList[[r]] <- pursuitMat ## just so i can analyze after
      
      
     # ROUND 1 ATTACKS #
      
      #convert pursuit mat into the attack probabilities
      attackProbs <- pursuitMat
      attackProbs <- ifelse(attackProbs == -1, probAttackEnemy, probAttackNon)
      
      #now agents decide whether to attack each other agent
      attackMat <- matrix(rbinom(popSize^2, 1, attackProbs), nrow = popSize)

  
     # ROUND 2 ATTACKS #
      
      #each agent looks at who their pursuits attacked/were attacked by, and complete triangle
      for(a in sample(1:popSize)){
        
        #complete attack triangles
        attackMat <- contingentAttack(a, attackMat, pursuitMat)
          
      }
      
      #once every agent did contingent attacks, rebinarize the matrix
      attackMat[attackMat == 2] <- 1
      
      print(mean(rowSums(attackMat, na.rm = T)))
        
      
      # GAME #
      
      # #get matrix of only reciprocated pursuits
      # friendsMatrix <- pursuitMat
      # friendsMatrix[friendsMatrix == -1] <- 0
      # friendsMatrix <- friendsMatrix * t(friendsMatrix)
      # diag(friendsMatrix) <- 0
      # 
      # 
      # for (gamer in sample(1:popSize)) {
      #   
      #   #pull indices of agents' reciprocated ties
      #   gamerGroup <- which(friendsMatrix[gamer, ] == 1)
      #   
      #   #add the gamer to the group indices
      #   gamerGroup <- c(gamerGroup, gamer)
      #   
      #   #subset the friends matrix to include only those in group
      #   gamersMat <- friendsMatrix[gamerGroup, gamerGroup, drop = F]
      #   
      #   #change the names back to original indices
      #   rownames(gamersMat) <- gamerGroup
      #   colnames(gamersMat) <- gamerGroup
      #   
      #   #Agents decide whether to play a PG game proportional to number of reciprocated ties in group
      #   #Game happens only if all agents agree
      #   if (nrow(gamersMat) == (numPursuits + 1) &
      #       prod(rbinom(
      #         nrow(gamersMat),
      #         1,
      #         rowSums(gamersMat, na.rm = T) / nrow(gamersMat)
      #       )) == 1) {
      #     
      #     #calculate the payoff each agent will receive
      #     agents <- payoff(gamersMat, gamerGroup, agents)
      #     
      #   }
      #   
      # }
      
      # UPDATE INIMITY # 

      #calculate enmity based on last round's direct attacks
      
      #bayesian updating when there was an attack
      num1 <- enemyMat * probAttackEnemy
      den1 <- num1 + (1 - enemyMat) * probAttackNon
      
      #bayesian updating when there was NOT an attack
      num0 <- enemyMat * (1 - probAttackEnemy)
      den0 <- num0 + (1 - enemyMat) * (1 - probAttackNon)
      
      #update the enemy matrix
      enemyMat <- ifelse(t(attackMat) == 1, num1 / den1, num0 / den0)
      
      #calculate enmity transitively
      
      #for each agent
      for(a in 1:popSize){
        
        #update enmity with others based on their contingent attacks
        enemyMat <- transUpdate(a, agents, pursuitMat, attackMat, enemyMat)
        
      }
      
      enemyMat <- round(enemyMat, digits = 5)
      
      enemyMatList[[r]] <- enemyMat
      
      
    } #round
    
    # SAVE DATA #
    
    # resultsList[[counter]] <- agents %>%
    #   mutate(loop = m, generation = g)
    # 
    # #just so i can see while it is running
    # print(
    #   agents %>%
    #     mutate(primTot = alliance + defense + displacement + defense) %>%
    #     count(primTot, name = "n")
    # )
    # 
    # counter <- counter + 1
    # 
    # agentsPrevious <- agents
    # 
    # # Reproduce #
    # agents <- reproduce(agents, selStrength, popSize)
    
    
  } #gen
  
} #loop

# results <- bind_rows(resultsList)
# results$pG <- probCoopGroup
# results$pN <- probCoopNon
# results$numRounds <- rounds
# results$numPursuits <- numPursuits
# results$flipProb <- flipProb
# 
# 
# #get the last loop number so it can be added to current
# if(file.exists(masterPath)){
#   existing <- read.csv(masterPath)
#   loop_offset <- max(existing$loop)
# } else {
#   loop_offset <- 0
# }
# 
# #add offset to loop number in new results
# results$loop <- results$loop + loop_offset
# 
# #append the results
# write.table(results, 
#             file = masterPath,
#             append = TRUE,
#             sep = ",",
#             col.names = !file.exists(masterPath),
#             row.names = FALSE)

```

```{r}

#see the average enemy prob across generations
eneMean <- data.frame("round" = 1:500,
                      "enemyProb" = NA)

for(i in 1:length(enemyMatList)){
  
  eneMean$enemyProb[i] <- mean(enemyMatList[[i]], na.rm = T)
  
  
}

table(round(enemyMat, digits = 1))

network <- graph_from_adjacency_matrix(friendsMatrix)

cliques(network, min = 2)







##############
networkMat <- enemyMat
networkMat <- round(networkMat, digits = 0)
#networkMat <- ifelse(networkMat == 1,0,1)
#friendsMatrix <- friendsMatrix * t(friendsMatrix)
diag(networkMat) <- 0

networkPlot <- graph_from_adjacency_matrix(networkMat, diag = F,  mode = "max")

plot(networkPlot, arrow.size = 0.2, width = 0.5)


##########




```


