---
title: "groupFormation"
author: "Clara Lavrador"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview

### Purpose


### Agents

Each agent will have a set of three traits and three preferences, which are randomly assigned from 0-100. 
They will also have a number of pursuits they can make per regrouping period.


### Life Cycle

1. Gather info
2. Pursue agents
3. Regroup

#### Gather info

Each agent will calculate three scores with respect to other agents: alliance value, reciprocity score, and overlap score.

The alliance value is calculated by taking the euclidean distance of the agent's preferences with the other agents' traits. 

The reciprocity score is the proportion of pursuits between the agent and each other agent that are reciprocated.

The overlap score is a measurement of pursuit similarity (both sent and received) between the agent and each other agent, with respect to third parties.
  It is based on David Pietraszewski's triadic primitives, whereby:
   
     alliance is the correlation between the agent and each other agents' sent pursuits (i.e., how much they pursue the same agents)
     
     generalization is the correlation between the agent and each other agents' received pursuits (i.e., how much they are pursued by the same agents)
     
     displacement is the correlation between the agent's sent pursuits and each other agents' received pursuits (i.e., how much the agents that agent b pursues pursue agent a)
     
     defense is the correlation between the agent's received and each other agents' sent pursuits (i.e., how much the agents that agent a is pursued by are pursued by agent b)

  
These three scores are then formulated to produce a single pursuit score:

  pursuit value = alliance value + (reciprocity score + overlap score)


#### Pursue agents

Each agent will pursue the agents with the highest pursuit values. 

Sent and received pursuits will be accumulated (and thus used to calculate overlap in the next reassessment period) 

#### Regroup

Agents who have reciprocated pursuits with each other will be considered "grouped" together for that reassessment period.

Then it all begins anew until the max number of reassessment periods has been reached.


### Analysis

For now I'm plotting the number of reciprocated edges over time across all model loops.
But I must think of a better representation... number of groups only tells half the story (bc they still differ in size)

Best thing would be to see if the model loops converge in some other aspect(s)

Also I could compare it to a model that does not include overlap score



## Model

### Packages 
```{r packages}

library(ggplot2)
library(igraph)
library(network)
library(networkDynamic)
library(intergraph)
library(ndtv)

```


### Parameters
```{r parameters}

#number of agents to generate
popSize <- 100

#pursuitRange <- c(1:5)

#number of pursuits each agent will have for each reassessment period
pursuitAmount <- 3

#number of times agents will reevaluate their pursuits
reassessments <- 300

#number of loops for the whole model
modelLoops <- 10

#to normalize euclidean distances, divide by highest possible euclidean distance considering vector length and range
maxEuc <- 174

```


### Functions
```{r functions}


#Agent Generate Function#

agentGenerate <- function(n){
  
  #randomly generate 3 traits and their corresponding preferences for each agent
  #alliance <- sample(c(1:100), n, replace = T)
  
  alliance <- abs(rnorm(popSize, mean = 50, sd = 10))
  
  alliance <- round(alliance, digits = 0)
  
  alliance <- (alliance/100)
  
  #put it all together in a df
  agents <- data.frame(alliance)
  
  return(agents)
  
}


weighted_jaccard <- function(x, y) {
  
  # Calculate the numerator: Sum of minimum intensities
  numerator <- sum(pmin(x, y))
  
  # Calculate the denominator: Sum of maximum intensities
  denominator <- sum(pmax(x, y))
  
  # Handle edge case where denominator is 0 (all zeros)
  if (denominator == 0) return(0)
  
  # Calculate weighted Jaccard similarity
  similarity <- numerator / denominator
  return(similarity)
}


#Overlap Score Function#
#calculates overlap between input agent and all other agents by comparing their sent and received pursuits; based on Dave's triadic primitives
#choose components to use in calculation

overlap <- function(agent, adjMat, components = c("alliance", "displacement", "generalization", "defense")) {
    
    #initialize component list
    scores <- list(
      alliance = NA,
      displacement = NA,
      defense = NA,
      generalization = NA
    )
    
    #Alliance#
    #calculate correlation of agent's row with other agents' rows  if ("alliance" %in% components) {
    if ("alliance" %in% components) {
      scores$alliance <- sapply(1:nrow(adjMat), function(i) {
        if (i == agent)
          return(NA)
        exclude_indices <- c(agent, i)
        x_filtered <- adjMat[i,-exclude_indices]
        agent_row_filtered <- adjMat[agent,-exclude_indices]
        weighted_jaccard(x_filtered, agent_row_filtered)
      })
    }
    
    #Displacement#
    #calculate correlation of agent's row with other agents' columns
    if ("displacement" %in% components) {
      scores$displacement <- sapply(1:ncol(adjMat), function(j) {
        if (j == agent)
          return(NA)
        exclude_indices <- c(agent, j)
        x_filtered <- adjMat[-exclude_indices, j]
        agent_filtered <- adjMat[agent,-exclude_indices]
        weighted_jaccard(x_filtered, agent_filtered)
      })
    }
    
    #Defense#
    #calculate correlation of agent's column with other agents' rows
    if ("defense" %in% components) {
      scores$defense <- sapply(1:nrow(adjMat), function(i) {
        if (i == agent)
          return(NA)
        exclude_indices <- c(agent, i)
        x_filtered <- adjMat[i,-exclude_indices]
        agent_filtered <- adjMat[-exclude_indices, agent]
        weighted_jaccard(x_filtered, agent_filtered)
      })
    }
    
    #Generalization#
    #calculate correlation of agent's column with other agents' columns
    if ("generalization" %in% components) {
      scores$generalization <- sapply(1:ncol(adjMat), function(j) {
        if (j == agent)
          return(NA)
        exclude_indices <- c(agent, j)
        x_filtered <- adjMat[-exclude_indices, j]
        agent_filtered <- adjMat[-exclude_indices, agent]
        weighted_jaccard(x_filtered, agent_filtered)
      })
    }
    
    #take average of only the selected components
    selected_scores <- do.call(cbind, scores[components])
    overlap_scores <- rowMeans(selected_scores, na.rm = TRUE)
    
    overlap_scores[agent] <- NA
    
    return(overlap_scores)
  }


#Reciprocity Function#

reciprocity <- function(agent, adjMat){
  
  #pull agent's column which represents their received pursuits
  received <- adjMat[,agent]
  
  #pull agent's row which represents their sent pursuits
  sent <- adjMat[agent,]
    
  #calculate proportions of pursuits which were received
  reciprocity_scores <- (received / (received + sent))
  
  #nullify their own reciprocity score
  reciprocity_scores[agent] <- NA
  
  return(reciprocity_scores)
  
}
  

#compare edges between two graphs

edge_compare <- function(graph1,graph2){
  
  sameEdges <- length(intersect(E(graph1),E(graph2)))
  
  return(sameEdges)
  
}



```


### Life Cycle

```{r lifeCycle}

start.time <- Sys.time()

#initialize results df
# results <- data.frame("loop" = rep(1:modelLoops,each = reassessments),
#                       "reassess" = rep(1:reassessments),
#                       "num_recip_edges" = NA)



#commenting out model loops for now
#for(m in 1:modelLoops) {
  
  #generate agents
  agents <- agentGenerate(popSize)

  #generate adjacency matrix (populate with 1)
  adjMat <- matrix(1, nrow = popSize, ncol = popSize)
  
  #store each reassessment period's cumulative adjacency matrix so that agents can reference the previous one when calculating their overlap/reciprocal scores
  matrix_list <- list(adjMat)
  
  #initialize list to store each reassessment's adjacency matrix as binary for that iteration so we can look at who was reciprocal at each reassessment period
  noncumMat_list <- list(adjMat)

  #for each reassessment... (agents look backwards so start from 2)
  for (t in 2:reassessments) {
    
    #create the empty noncumulative adjacency matrix
    noncumMat <- matrix(0, nrow = popSize, ncol = popSize)
    
    #for each agent...
    for (a in 1:nrow(agents)) {
      
      #calculate the overlap scores between themselves and the other agents
      overlap_scores <- overlap(a, matrix_list[[t - 1]], c("alliance", "generalization", "displacement"))
      
      #calculate the reciprocity scores between themselves and the other agents
      reciprocity_scores <- reciprocity(a, matrix_list[[t - 1]])
      
      #calculate the alliance values between their prefs and the other agents' traits
      alliance_values <- agents$alliance
      
      #Combine overlap, reciprocity, and alliance scores into single calculation
      pursuit_scores <- overlap_scores * reciprocity_scores * alliance_values
      
      #Randomize number of pursuits for the agent
      #pursuitAmount <- sample(pursuitRange, 1, replace = T)
      
      #sort by product in descending order and get the indices of the top value agents
      pursuit_indices <- order(-pursuit_scores)[1:pursuitAmount]
      
      #add 1 to their pursuits in the cumulative the adjacency matrix
      adjMat[a, pursuit_indices] <- adjMat[a, pursuit_indices] + 1
      
      #and to the individual one
      noncumMat[a, pursuit_indices] <- 1
      
      
    }
    
    #add the adjMat to the matrix list
    matrix_list[[t]] <- adjMat
    
    #add the non cumulative matrix to its list
    noncumMat_list[[t]] <- noncumMat
    
  }
  
  
  #convert the noncum adjacency matrices into igraph networks
  # igraph_list <- lapply(noncumMat_list, graph_from_adjacency_matrix)
  # 
  # #loop through each igraph object
  # for (g in 1:length(igraph_list)) {
  #   
  #   #only include edges that are reciprocated
  #   igraph_list[[g]] <- subgraph.edges(igraph_list[[g]],
  #                                             E(igraph_list[[g]])[which_mutual(igraph_list[[g]])],
  #                                             delete.vertices = F) #maybe i can just do this with true to get groups
  #   
  #   #add number of reciprocal edges to results df
  #   print(length(subgraph.edges(igraph_list[[g]], E(igraph_list[[g]])[which_mutual(igraph_list[[g]])])))
  # 
  #   
    
 # }
#}

#the first reassess from each loop displays recip edge number as popSize because of intial matrix set up, so fix to 0
#results$num_recip_edges[results$reassess == 1] <- 0

end.time <- Sys.time()

print(end.time - start.time)

```

### Analysis

```{r analysis}


### Igraph Network

## convert to igraph
#convert the noncum adjacency matrices into igraph networks to see groups over time
#deleting non recip edges instead of denoting them makes dynamic net run much faster

igraph_list <- lapply(noncumMat_list, graph_from_adjacency_matrix, add.rownames = T)
  
#loop through each igraph object
for (g in 1:length(igraph_list)) {
  
  #only include edges that are reciprocated
  igraph_list[[g]] <- subgraph.edges(igraph_list[[g]],
                                     E(igraph_list[[g]])[which_mutual(igraph_list[[g]])],
                                     delete.vertices = F) #maybe i can just do this with true to get groups
  
  #just curious
  print(length(subgraph.edges(igraph_list[[g]], E(igraph_list[[g]])[which_mutual(igraph_list[[g]])])))
  
  
}

#add vertex attributes
for(r in 1:length(igraph_list)) {
  
  V(igraph_list[[r]])$name <- c(1:100)
  
  V(igraph_list[[r]])$alliance <- agents$alliance[match(V(igraph_list[[r]])$name, rownames(agents))]
  
}

#Optional igraph plot (just change index)
plot(all_igraph_list[[300]],
     layout = layout.fruchterman.reingold,
     #vertex.label = NA,
     #  rescale = F,
     layout = layout_with_kk,
     vertex.size = 8,
     #vertex.color = vertex.color,
     vertex.label.color = "black",
     vertex.shape = "circle",
     edge.arrow.size = 0.3)


## Clique data
#cliques differ from groups in that the subgraphs must be interconnected to be counted as clique

clique_results <- data.frame("reassessment" = rep(1:reassessments, each = 4),
                             "percentInClique" = NA, #percent of agents in cliques of 3+
                             "cliqueSize" = rep(1:4), #clique sizes 1-4
                             "percentOfSize" = NA)

#for each reassessment period...
for (r in 1:length(igraph_list)) {
  
  #...get percent of agents in a clique of at least 3
  percentInClique <- length(unique(unlist(max_cliques(igraph_list[[r]], min = 3))))/popSize
  
  #get total number of cliques of 1,2,3,4 to serve as denominator
  #total_cliques[r] <- count_max_cliques(igraph_list[[r]], min = 1)
  
  #get number of cliques of each size
  cliques_two <- length(unique(unlist(max_cliques(igraph_list[[r]], min = 2, max = 2))))/popSize
  cliques_three <- length(unique(unlist(max_cliques(igraph_list[[r]], min = 3, max = 3))))/popSize
  cliques_four <- length(unique(unlist(max_cliques(igraph_list[[r]], min = 4, max = 4))))/popSize
  isolates <- length(unique(unlist(max_cliques(igraph_list[[r]], min = 1, max = 1))))/popSize
  
  #get number of constant edges between each reassessment
  constant_edges <- edge_compare(igraph_list[[r]], igraph_list[[r+1]])

  
  #add to results df
  clique_results$percentInClique[clique_results$reassess == r] <- percentInClique
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 1] <- isolates
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 2] <- cliques_two
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 3] <- cliques_three
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 4] <- cliques_four
  clique_results$constant_edges[clique_results$reassessment == r+1] <- constant_edges
  
}

clique_results$percentInClique[clique_results$reassessment == 1] <- 0
clique_results$percentOfSize[clique_results$reassessment == 1 & clique_results$cliqueSize == 1] <- popSize/100



#plot to see percent of agents that are part of a clique (3+ agents) across regrouping periods
ggplot(data = clique_results,
       aes(x = reassessment,
           y = percentInClique)) +
  labs(title = "Percent of agents in cliques per reassessment",
       subtitle = "reciprocity(0-1) * overlap(0-1) * alliance(0-1)",
       x = "reassessment period",
       y = "percent of agents in a clique (3+)") +
  geom_line()


#plot to see which edges remain constant over reassessment periods
ggplot(data = clique_results,
       aes(x = reassessment,
           y = constant_edges)) +
  labs(title = "Edge Constancy ",
       x = "regrouping period",
       y = "number of retained edges from previous regrouping") +
  geom_smooth()


#plot to see clique size distribution across regrouping periods
#still trying to reword but basically, I am plotting the percent of agents whose max clique size is each of the clique sizes
ggplot(data = clique_results,
       aes(x = reassessment,
           y = percentOfSize,
           color = as.factor(cliqueSize))) +
  labs(title = "Maximal Clique",
       subtitle = "percent of agents whose maximal clique is of size 'clique size'",
       x = "reassessment period",
       y = "percent of agents in maximal clique",
       color = "clique size") +
  geom_smooth()


#number of isolates

isolate_list <- c(rep(NA, 350))
for(r in 1:length(igraph_list)){
  
    isolates <- length(unique(unlist(max_cliques(igraph_list[[r]], min = 1, max = 1))))
    isolate_list[r] <- isolates
  
}


### Dynamic Network
#we shall create dynamic network of one model loop

start.time <- Sys.time()

#Create Dynamic Network#
#convert the igraph to network objects so we can make dynamic net
networkList <- lapply(igraph_list, asNetwork)

#create dynamic network from network list
dynamicNet <- networkDynamic(network.list = networkList, 
                                 vertex.pid = "vertex.names",
                                 create.TEAs = T,)

#Vertex Attributes#
#add alliance value and corresponding alliance value color as vertex attributes

#create vector of unique alliance values in order 
allianceVals <- sort(unique(agents$alliance))

#create color ramp palette
cols <- setNames(colorRampPalette(c("red", "yellow", "green"))(length(allianceVals)), allianceVals)

# initialize color vector
vertColors <- rep(NA, popSize)

# assign colors based on alliance values
for (i in 1:nrow(agents)) {
  
  a <- agents$alliance[i]
  
  vertColors[i] <- cols[as.character(a)]
}


cbind(network.vertex.names(dynamicNet), rownames(agents)) #ensure they are the same

#set the attributes
network::set.vertex.attribute(dynamicNet, attrname = "alliance", value = agents$alliance, v = seq_along(network.vertex.names(dynamicNet))) 
network::set.vertex.attribute(dynamicNet, attrname = "color", value = vertColors, v = seq_along(network.vertex.names(dynamicNet))) 


#Render Movie#
render.d3movie(dynamicNet, filename = "dynNet.html", vertex.col = dynamicNet %v% "color")

end.time <- Sys.time()

print(end.time - start.time)


```

```{r network}

#Create network list
network_list <- lapply(noncumMat_list, as.network.matrix, add.rownames = T)



#Dynamic Network
#we shall create dynamic network of one model loop

start.time <- Sys.time()

#Create Dynamic Network#

#create dynamic network from network list
dynamicNet <- networkDynamic(network.list = network_list, 
                                 vertex.pid = "vertex.names",
                                 create.TEAs = F)

#Vertex Attributes#
#add alliance value and corresponding alliance value color as vertex attributes

#create vector of unique alliance values in order 
allianceVals <- sort(unique(agents$alliance))

#create color ramp palette
cols <- setNames(colorRampPalette(c("red", "yellow", "green"))(length(allianceVals)), allianceVals)

# initialize color vector
vertColors <- rep(NA, popSize)

# assign colors based on alliance values
for (i in 1:nrow(agents)) {
  
  a <- agents$alliance[i]
  
  vertColors[i] <- cols[as.character(a)]
}


cbind(network.vertex.names(dynamicNet), rownames(agents)) #ensure they are the same

#set the attributes
network::set.vertex.attribute(dynamicNet, attrname = "alliance", value = agents$alliance, v = seq_along(network.vertex.names(dynamicNet))) 
network::set.vertex.attribute(dynamicNet, attrname = "color", value = vertColors, v = seq_along(network.vertex.names(dynamicNet))) 


#Render Movie#
render.d3movie(dynamicNet, filename = "dynNet.html", vertex.col = dynamicNet %v% "color")

end.time <- Sys.time()

print(end.time - start.time)



```














```{r trash}

for (r in 1:length(igraph_list)) {
    

    prop1 <- length(igraph::components(igraph_list[[r]])$csize[igraph::components(igraph_list[[r]])$csize == 1])/igraph::components(igraph_list[[r]])$no
    
    prop2 <- length(igraph::components(igraph_list[[r]])$csize[igraph::components(igraph_list[[r]])$csize == 2])/igraph::components(igraph_list[[r]])$no
    
    prop3 <- length(igraph::components(igraph_list[[r]])$csize[igraph::components(igraph_list[[r]])$csize == 3])/igraph::components(igraph_list[[r]])$no
    
    prop4 <- length(igraph::components(igraph_list[[r]])$csize[igraph::components(igraph_list[[r]])$csize == 4])/igraph::components(igraph_list[[r]])$no
    
    prop5 <- length(igraph::components(igraph_list[[r]])$csize[igraph::components(igraph_list[[r]])$csize == 5])/igraph::components(igraph_list[[r]])$no
    
    
    #get number of closed triangles
    max_cliques(igraph_list[[r]], min = 3)
  
    #add number of reciprocal edges to results df
    igraph_results$proportion[igraph_results$reassess == r & igraph_results$groupSize == 1] <- prop1
    igraph_results$proportion[igraph_results$reassess == r & igraph_results$groupSize == 2] <- prop2
    igraph_results$proportion[igraph_results$reassess == r & igraph_results$groupSize == 3] <- prop3
    igraph_results$proportion[igraph_results$reassess == r & igraph_results$groupSize == 4] <- prop4
    igraph_results$proportion[igraph_results$reassess == r & igraph_results$groupSize == 5] <- prop5

}

 #Defense#
      #calculate correlation of agent's column with other agents' rows
      
      defense <- apply(adjMat, 1, function(x){
        
        
        
        weighted_jaccard(x, adjMat[, agent])
      })
      
      
## Plot group size
# plot the prevalence of different sized groups over time (reassessment periods)

igraph_results <- data.frame(
                      "reassess" = rep(1:reassessments, each = 5),
                      "groupSize" = rep(1:5),
                      "proportion" = NA)


   
#plot it
ggplot(data = igraph_results,
       aes(x = reassess,
           y = proportion,
           color = as.factor(groupSize))) +
  labs(title = "Prevalence of different sized groups over reassesments",
       subtitle = "(reciprocity(0-1) * overlap(0-1)) * alliance(0-1)",
       x = "reassessment period",
       y = "percent of groups of size x",
       color = "group size") +
  geom_line()


## single plot syntax
# you can plot any for fun

V(igraph_list[[350]])$alliance <- agents$alliance[match(V(igraph_list[[350]])$name, rownames(agents))]

V(igraph_list[[350]])$color <- V(igraph_list[[350]])$

  


for (r in 1:length(igraph_list)) {
  
  #...get percent of agents in a clique of at least 3
  percentInClique <- length(unique(unlist(max_cliques(igraph_list[[r]], min = 3))))/popSize
  
  #get total number of cliques of 1,2,3,4 to serve as denominator
  total_cliques[r] <- count_max_cliques(igraph_list[[r]], min = 1)
  
  #get number of cliques of each size
  cliques_two <- count_max_cliques(igraph_list[[r]], min = 2, max = 2)/total_cliques[r]
  cliques_three <- count_max_cliques(igraph_list[[r]], min = 3, max = 3)/total_cliques[r]
  cliques_four <- count_max_cliques(igraph_list[[r]], min = 4, max = 4)/total_cliques[r]
  isolates <- count_max_cliques(igraph_list[[r]], min = 1, max = 1)/total_cliques[r]

  #add to results df
  clique_results$percentInClique[clique_results$reassess == r] <- percentInClique
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 1] <- isolates
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 2] <- cliques_two
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 3] <- cliques_three
  clique_results$percentOfSize[clique_results$reassess == r & clique_results$cliqueSize == 4] <- cliques_four
  
}



#Testing Space#
adjMatTest <- matrix(c(1, 5, 2, 3,
                   1, 1, 4, 5,
                   2, 4, 6, 3,
                   3, 5, 6, 6), 
                 nrow = 4, byrow = TRUE)


agent <- 2
i <- 3

exclude_indices <- c(agent, i)
x_filtered <- adjMatTest[i, -exclude_indices]
agent_row_filtered <- adjMatTest[agent, -exclude_indices]
      
weighted_jaccard(x_filtered, agent_row_filtered)



x_filtered <- adjMatTest[-exclude_indices, i]
agent_row_filtered <- adjMatTest[-exclude_indices, agent]
      
weighted_jaccard(x_filtered, agent_row_filtered)



overlap(2, adjMatTest, c("alliance","generalization"))









```









