---
title: "groupFormation"
author: "Clara Lavrador"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview

### Purpose


### Agents

Each agent will have a set of three traits and three preferences, which are randomly assigned from 0-100. 
They will also have a number of pursuits they can make per regrouping period.


### Life Cycle

1. Gather info
2. Pursue agents
3. Regroup

#### Gather info

Each agent will calculate three scores with respect to other agents: alliance value, reciprocity score, and overlap score.

The alliance value is calculated by taking the euclidean distance of the agent's preferences with the other agents' traits. 

The reciprocity score is the proportion of pursuits between the agent and each other agent that are reciprocated.

The overlap score is a measurement of pursuit similarity (both sent and received) between the agent and each other agent, with respect to third parties.
  It is based on David Pietraszewski's triadic primitives, whereby:
   
     alliance is the correlation between the agent and each other agents' sent pursuits (i.e., how much they pursue the same agents)
     
     generalization is the correlation between the agent and each other agents' received pursuits (i.e., how much they are pursued by the same agents)
     
     displacement is the correlation between the agent's sent pursuits and each other agents' received pursuits (i.e., how much the agents that agent b pursues pursue agent a)
     
     defense is the correlation between the agent's received and each other agents' sent pursuits (i.e., how much the agents that agent a is pursued by are pursued by agent b)

  
These three scores are then formulated to produce a single pursuit score:

  pursuit value = alliance value + (reciprocity score + overlap score)


#### Pursue agents

Each agent will pursue the agents with the highest pursuit values. 

Sent and received pursuits will be accumulated (and thus used to calculate overlap in the next reassessment period) 

#### Regroup

Agents who have reciprocated pursuits with each other will be considered "grouped" together for that reassessment period.

Then it all begins anew until the max number of reassessment periods has been reached.


### Analysis

For now I'm plotting the number of reciprocated edges over time across all model loops.
But I must think of a better representation... number of groups only tells half the story (bc they still differ in size)

Best thing would be to see if the model loops converge in some other aspect(s)

Also I could compare it to a model that does not include overlap score



## Model

### Packages 
```{r packages}

library(ggplot2)
library(igraph)
library(network)
library(networkDynamic)
library(intergraph)
library(ndtv)

```


### Parameters
```{r parameters}

#number of agents to generate
popSize <- 100

#pursuitRange <- c(1:5)

#number of pursuits each agent will have for each reassessment period
pursuitAmount <- 3

#number of times agents will reevaluate their pursuits
reassessments <- 500

#number of loops for the whole model
modelLoops <- 10

#to normalize euclidean distances, divide by highest possible euclidean distance considering vector length and range
maxEuc <- 174

all_primitives <- c("alliance", "displacement", "defense", "generalization")

all_combos <- unlist(
  lapply(1:length(all_primitives), function(k) {
    combn(all_primitives, k, simplify = FALSE)
  }),
  recursive = FALSE
)

prim_abbrev <- c("A",
"P",
"D",
"G",
"AP",
"AD",
"AG",
"PD",
"PG",
"DG",
"APD",
"APG",
"ADG",
"PDG",
"APDG")

primies <- as.data.frame(prim_abbrev)

#conditions to test



```


### Functions
```{r functions}


#Agent Generate Function#

agentGenerate <- function(n){
  
  #randomly generate 3 traits and their corresponding preferences for each agent

  alliance <- abs(rnorm(popSize, mean = 50, sd = 10))
  
  alliance <- round(alliance, digits = 0)
  
  alliance <- (alliance/100)
  
  pursuit_amount <- sample(2:5, n, replace = T)
  
  #put it all together in a df
  agents <- data.frame(alliance, pursuit_amount)
  
  return(agents)
  
}


weighted_jaccard <- function(x, y) {
  
  # Calculate the numerator: Sum of minimum intensities
  numerator <- sum(pmin(x, y))
  
  # Calculate the denominator: Sum of maximum intensities
  denominator <- sum(pmax(x, y))
  
  # Handle edge case where denominator is 0 (all zeros)
  if (denominator == 0) return(0)
  
  # Calculate weighted Jaccard similarity
  similarity <- numerator / denominator
  return(similarity)
}


#Overlap Score Function#
#calculates overlap between input agent and all other agents by comparing their sent and received pursuits; based on Dave's triadic primitives
#choose components to use in calculation

overlap <- function(agent, adjMat, component) {
  
  scores <- switch(component,
    

    #Alliance#
    #calculate correlation of agent's row with other agents' rows 
      "alliance" = sapply(1:nrow(adjMat), function(i) {
        if (i <= agent)
          return(NA)
        exclude_indices <- c(agent, i)
        x_filtered <- adjMat[i,-exclude_indices]
        agent_row_filtered <- adjMat[agent,-exclude_indices]
        weighted_jaccard(x_filtered, agent_row_filtered)
      }),
    
    #Displacement#
    #calculate correlation of agent's row with other agents' columns
      "displacement" = sapply(1:nrow(adjMat), function(j) {
        if (j <= agent)
          return(NA)
        exclude_indices <- c(agent, j)
        x_filtered <- adjMat[-exclude_indices, j]
        agent_filtered <- adjMat[agent,-exclude_indices]
        weighted_jaccard(x_filtered, agent_filtered)
      }),
    
    #Defense#
    #calculate correlation of agent's column with other agents' rows
      "defense" = sapply(1:ncol(adjMat), function(i) {
        if (i <= agent)
          return(NA)
        exclude_indices <- c(agent, i)
        x_filtered <- adjMat[i,-exclude_indices]
        agent_filtered <- adjMat[-exclude_indices, agent]
        weighted_jaccard(x_filtered, agent_filtered)
      }),
    
    #Generalization#
    #calculate correlation of agent's column with other agents' columns
      "generalization" = sapply(1:ncol(adjMat), function(j) {
        if (j <= agent)
          return(NA)
        exclude_indices <- c(agent, j)
        x_filtered <- adjMat[-exclude_indices, j]
        agent_filtered <- adjMat[-exclude_indices, agent]
        weighted_jaccard(x_filtered, agent_filtered)
      })
  )

    return(scores)
  
  }



#Reciprocity Function#

reciprocity <- function(agent, adjMat){
  
  #pull agent's column which represents their received pursuits
  received <- adjMat[,agent]
  
  #pull agent's row which represents their sent pursuits
  sent <- adjMat[agent,]
    
  #calculate proportions of pursuits which were received
  reciprocity_scores <- (received / (received + sent))
  
  #nullify their own reciprocity score
  reciprocity_scores[agent] <- NA
  
  return(reciprocity_scores)
  
}
  


```


### Life Cycle

```{r lifeCycle}

start.time <- Sys.time()

#initialize results df
results <- data.frame("combo" = rep(1:length(all_combos),each = reassessments),
                      "reassess" = rep(1:reassessments),
                      "num_isolates" = NA,
                      "num_groups" = NA,
                      "num_four" = NA,
                      "assort_alliance" = NA)

lastGraphs <- list()

#Generate agents
agents <- agentGenerate(popSize)

#parallel process here

#run for all possible primitive combos
for(c in 1:length(all_combos)) {
  
  primitives <- all_combos[[c]]
  
  #generate adjacency matrix (populate with 1)
  adjMat <- matrix(1, nrow = popSize, ncol = popSize)
  
  #store each reassessment period's cumulative adjacency matrix so that agents can reference the previous one when calculating their overlap/reciprocal scores
  matrix_list <- list(adjMat)
  
  #initialize list to store each reassessment's adjacency matrix as binary for that iteration so we can look at who was reciprocal at each reassessment period
  noncumMat_list <- list(adjMat)
  

  #for each reassessment... (agents look backwards so start from 2)
  for (t in 2:reassessments) {
    
    #create the empty noncumulative adjacency matrix
    noncumMat <- matrix(0, nrow = popSize, ncol = popSize)
    
    #Calculate overlap in previous adjacency matrix
    #make empty matrix list
    overlap_matrices <- list()
    
    #for each primitive you will be using...
    for (prim in primitives) {
      
      #create a matrix for that primitive
      overlap_matrices[[prim]] <- matrix(NA, nrow = popSize, ncol = popSize)
      
      #use overlap function to calculate overlap for that primitive
      for (r in 1:nrow(overlap_matrices[[prim]])) {
        overlap_matrices[[prim]][r,] <- overlap(r, matrix_list[[t-1]], prim)
      }
      
      #fold the the matrix
      overlap_matrices[[prim]][lower.tri(overlap_matrices[[prim]])] <- t(overlap_matrices[[prim]])[lower.tri(overlap_matrices[[prim]])]
    }
    


    # Calculate the final overlap score matrix
    final_overlap_matrix <- Reduce("+", lapply(primitives, function(prim) overlap_matrices[[prim]])) / length(primitives)

    
 #Loop through each agent to compute their pursuits
    
    #for each agent...
    for (a in 1:nrow(agents)) {
      
      #pull the overlap scores between themselves and the other agents from the final overlap matrix
      overlap_scores <- final_overlap_matrix[a,]
      
      #calculate the reciprocity scores between themselves and the other agents
      reciprocity_scores <- reciprocity(a, matrix_list[[t - 1]])
      
      #calculate the alliance values between their prefs and the other agents' traits
      alliance_values <- agents$alliance
      
      #Combine overlap, reciprocity, and alliance scores into single calculation
      pursuit_scores <- overlap_scores * reciprocity_scores * alliance_values
      
      #Randomize number of pursuits for the agent
      pursuitAmount <- pursuitAmount
      
      #sort by product in descending order and get the indices of the top value agents
      pursuit_indices <- order(-pursuit_scores)[1:pursuitAmount]
      
      #add 1 to their pursuits in the cumulative the adjacency matrix
      adjMat[a, pursuit_indices] <- adjMat[a, pursuit_indices] + 1
      
      #and to the individual one
      noncumMat[a, pursuit_indices] <- 1
      
      
    }
    
    #add the adjMat to the matrix list
    matrix_list[[t]] <- adjMat
    
    #add the non cumulative matrix to its list
    noncumMat_list[[t]] <- noncumMat
    
  }
  
  
  igraph_list <-lapply(noncumMat_list, graph_from_adjacency_matrix, add.rownames = T)
  
  #loop through each igraph object
  for (g in 1:length(igraph_list)) {
    
    #only include edges that are reciprocated
    igraph_list[[g]] <- subgraph.edges(igraph_list[[g]],
                                       E(igraph_list[[g]])[which_mutual(igraph_list[[g]])],
                                       delete.vertices = F) #maybe i can just do this with true to get groups
    
    #get total number of edges
    numEdges <- length(E(igraph_list[[g]]))
    
    
    #get edge overlap with previous graph
    intersectLast <- ifelse(g > 1, length(E(intersection(igraph_list[[g-1]],igraph_list[[g]]))), NA)
    
    #get number of groups
    num_groups <- igraph::components(igraph_list[[g]])$no
    
    #get number of isolates
    num_isolates <- length(max_cliques(igraph_list[[g]], min = 1, max = 1))
      
    #get number of closed groups of 4
    num_four <- length(max_cliques(igraph_list[[g]], min = 4, max = 4))
    
    V(igraph_list[[g]])$name <- c(1:100)
  
    V(igraph_list[[g]])$alliance <- agents$alliance[match(V(igraph_list[[g]])$name, rownames(agents))]
  
    #add number of reciprocal edges to results df
    results$numEdges[results$reassess == g & results$combo == c] <- numEdges
    results$intersectLast[results$reassess == g & results$combo == c] <- intersectLast
    results$num_groups[results$reassess == g & results$combo == c] <- num_groups
    results$num_isolates[results$reassess == g & results$combo == c] <- num_isolates
    results$num_four[results$reassess == g & results$combo == c] <- num_four
    results$assort_alliance[results$reassess == g & results$combo == c] <- assortativity(igraph_list[[g]], V(igraph_list[[g]])$alliance)


  }
  
  
  lastGraphs[[c]] <- igraph_list[[reassessments]]
  
  print(c)

  
} 
  

results$numEdges[results$reassess == 1] <- 0
results$intersectLast[results$reassess == 2] <- 0
results$num_isolates[results$reassess == 1] <- 100

results$prims <- primies$prim_abbrev[match(results$combo, rownames(primies))]





end.time <- Sys.time()

print(end.time - start.time)


```

### Analysis

```{r analysis}



write.csv(results, "C:/Users/clara/Downloads/groupFormation/comboRes10x.csv")
write.csv(agents, "C:/Users/clara/Downloads/groupFormation/agents10.csv")



ggplot(data = results,
       aes(x = reassess,
           y = numEdges,
           color = as.factor(prims))) +
  labs(title = "",
       subtitle = "",
       x = "reassessment period",
       y = "percent of edges that are reciprocated",
       color = "primitives") +
  geom_smooth(se= F)


```

















