---
title: "groupFormation"
author: "Clara Lavrador"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview

### Purpose


### Agents

Each agent will have a set of three traits and three preferences, which are randomly assigned from 0-100. 
They will also have a number of pursuits they can make per regrouping period.


### Life Cycle

1. Gather info
2. Pursue agents
3. Regroup

#### Gather info

Each agent will calculate three scores with respect to other agents: association value, reciprocity score, and overlap score.

The association value is calculated by taking the euclidean distance of the agent's preferences with the other agents' traits. 

The reciprocity score is the proportion of pursuits between the agent and each other agent that are reciprocated.

The overlap score is a measurement of pursuit similarity (both sent and received) between the agent and each other agent, with respect to third parties.
  It is based on David Pietraszewski's triadic primitives, whereby:
   
     alliance is the correlation between the agent and each other agents' sent pursuits (i.e., how much they pursue the same agents)
     
     generalization is the correlation between the agent and each other agents' received pursuits (i.e., how much they are pursued by the same agents)
     
     displacement is the correlation between the agent's sent pursuits and each other agents' received pursuits (i.e., how much the agents that agent b pursues pursue agent a)
     
     defense is the correlation between the agent's received and each other agents' sent pursuits (i.e., how much the agents that agent a is pursued by are pursued by agent b)

  
These three scores are then formulated to produce a single pursuit score:

  pursuit value = association value + (reciprocity score + overlap score)


#### Pursue agents

Each agent will pursue the agents with the highest pursuit values. 

Sent and received pursuits will be accumulated (and thus used to calculate overlap in the next reassessment period) 

#### Regroup

Agents who have reciprocated pursuits with each other will be considered "grouped" together for that reassessment period.

Then it all begins anew until the max number of reassessment periods has been reached.


### Analysis

For now I'm plotting the number of reciprocated edges over time across all model loops.
But I must think of a better representation... number of groups only tells half the story (bc they still differ in size)

Best thing would be to see if the model loops converge in some other aspect(s)

Also I could compare it to a model that does not include overlap score



## Model

### Packages 
```{r packages}

library(ggplot2)
library(igraph)
library(network)
library(networkDynamic)
library(intergraph)
library(ndtv)
library(dplyr)
library(doParallel)

```


### Parameters
```{r parameters}

#number of agents to generate
popSize <- 100

#number of pursuits each agent will have for each reassessment period
pursuitAmount <- 3

#number of times agents will reevaluate their pursuits
reassessments <- 200

#number of loops for the whole model
modelLoops <- 2

#number of generations
generations <- 50

#SD of association value of agents in the eyes of other agents
AV_SD <- 0.05

#set strength of selection
selStrength <- 6

#multiplier
multiplier <- 1.2

#mutation probability that binary prim trait will turn on/off 
flipProb <- 0.01

myPath <- "C:/Users/clavra/Documents/groupFormation/resultsMain.csv"



```


### Functions
```{r functions}

#Agent Generate Function#

agentGenerate <- function(n){
  
  #generate IDs for each agent
  #ID <- 1:n
  
  #generate association value
  association <- abs(rnorm(n, mean = 50, sd = 10))
  association <- round(association, digits = 0)
  association <- (association/100)
  
  #generate weight of each prim for each agent 
  A <- rbinom(n, 1, 0.5)
  
  G <- rbinom(n, 1, 0.5)
  
  P <- rbinom(n, 1, 0.5)
  
  D <- rbinom(n, 1, 0.5)
  
  payoff <- 0
  
  #put it all together in a df
  agents <- data.frame(association, A, G, P, D, payoff)
    
  
  return(agents)
  
}


#Association Matrix#
#each agent will see each other agent's association value slightly differently

makeAssociationMat <- function(agentDF, sd, n){
  
  associationMat <- matrix(, nrow = n, ncol = n)
  
  for(a in 1:nrow(agentDF)){
    
    associationMat[,a] <-  rnorm(n, mean = agentDF$association[a], sd = sd)
    
  }
  
  diag(associationMat) <- NA
  
  return(associationMat)
  
}
  
  
#Weighted Jaccard#
#used to calculate overlap#

weighted_jaccard <- function(x, y) {
  
  # Calculate the numerator: Sum of minimum intensities
  numerator <- sum(pmin(x, y))
  
  # Calculate the denominator: Sum of maximum intensities
  denominator <- sum(pmax(x, y))
  
  # Handle edge case where denominator is 0 (all zeros)
  if (denominator == 0) return(0)
  
  # Calculate weighted Jaccard similarity
  similarity <- numerator / denominator
  return(similarity)
}



#New Overlap Function#
#each agent will calculate their alliance, gen, displacement, and defense overlap between themselves and every other agent
#this vector will be stored in the row 

#fold triangles for alliance and gen, transpose for displacement to defense

#output = 4 matrices

overlap_prims<- function(adjMat) {
  popSize <- nrow(adjMat)  
  
  #Make empty matrices to store overlap for each prim
  allianceMat <- matrix(0, nrow = popSize, ncol = popSize)
  generalizationMat <- matrix(0, nrow = popSize, ncol = popSize)
  displacementMat <- matrix(0, nrow = popSize, ncol = popSize)
  defenseMat <- matrix(0, nrow = popSize, ncol = popSize)
  
  #loop through each agent
  for (agent in 1:popSize) {
    
    #store agent's row and column
    agent_row <- adjMat[agent, ]
    agent_col <- adjMat[, agent]
    
    ## Alliance (row-row) ##
    alliance <- rep(NA_real_, popSize)
    
    if (agent < popSize) {
      for (i in (agent + 1):popSize) {
        exclude_indices <- c(agent, i)
        agent_filtered <- agent_row[-exclude_indices]
        other_filtered <- adjMat[i, -exclude_indices]
        alliance[i] <- weighted_jaccard(agent_filtered, other_filtered)
      }
    }
    allianceMat[agent, ] <- alliance
    
    
    ## Generalization (col-col) ##
    generalization <- rep(NA_real_, popSize)
    
    if (agent < popSize) {
      for (j in (agent + 1):popSize) {
        exclude_indices <- c(agent, j)
        agent_filtered <- agent_col[-exclude_indices]
        other_filtered <- adjMat[-exclude_indices, j]
        generalization[j] <- weighted_jaccard(agent_filtered, other_filtered)
      }
    }
    generalizationMat[agent, ] <- generalization
    
    
    ## Displacement (row-col) ##
    displacement <- rep(NA_real_, popSize)
    
    for (j in 1:popSize) {
      if (j == agent) next
      exclude_indices <- c(agent, j)
      agent_filtered <- agent_row[-exclude_indices]
      other_filtered <- adjMat[-exclude_indices, j]
      displacement[j] <- weighted_jaccard(agent_filtered, other_filtered)
    }
    
    displacementMat[agent, ] <- displacement
  }
  
  ## Complete the Matrices ##
  
  # Symmetric matrices
  allianceMat[lower.tri(allianceMat)] <- t(allianceMat)[lower.tri(allianceMat)]
  generalizationMat[lower.tri(generalizationMat)] <- t(generalizationMat)[lower.tri(generalizationMat)]
  
  # Displacement + transpose = defense
  defenseMat <- t(displacementMat)
  
  return(list(
    A = allianceMat,
    G = generalizationMat,
    P = displacementMat,
    D = defenseMat
  ))
}


#Overlap Score Calculation#
#average based on agent's prim strat

overlap_mean <- function(agent, agentDF, matrixList){
  
  #for a given agent:
  
  #get her binary prim vector from agent DF
  prims <- agentDF[agent, c("A","G","P","D")]
  
  #get the names of the prims she is to use
  selectedPrims <- names(prims)[prims == 1]
  
  #get only the prim matrices she will be using
  selectedMats <- matrixList[selectedPrims]
    
  #get her rows from the prim mats and bind them together
  row_values <- do.call(rbind, lapply(selectedMats, function(mat)
        mat[agent,]))
    
  #take the element wise average
  means <- if(is.null(row_values)) rep(1, popSize) else colMeans(row_values)
    
  #return the vector of means, where each element is her overlap score with each other agent
  return(means)
  
}




#Reciprocity Function#

reciprocity <- function(agent, adjMat){
  
  #pull agent's column which represents their received pursuits
  received <- adjMat[,agent]
  
  #pull agent's row which represents their sent pursuits
  sent <- adjMat[agent,]
    
  #calculate proportions of pursuits which were received
  reciprocity_scores <- (received / (received + sent))
  
  #nullify their own reciprocity score
  reciprocity_scores[agent] <- NA
  
  return(reciprocity_scores)
  
}


#Contributions#

payoff <- function(gameMatrix, gamers, agentDF){
  
  #start the sum
  contributionTot <- 0
  
  #for each agent...
  for(agent in gamers){
    
    #their max contribution is their association value
    maxContribution <- agentDF$association[row.names(agentDF) == agent] * 100
    
    #get the percent of the AV they will contribute, which is proportional to number of recip ties within group
    proportionTied <- if (nrow(gameMatrix) > 1) sum(gameMatrix[as.character(agent), ])/(nrow(gameMatrix)-1) else 0
    
    
    #calculate their contribution 
    contribution <- maxContribution * proportionTied
    
    #add their total contribution to those of the other agents
    contributionTot <- contributionTot + contribution
    
    #add their noncontributed pay to their corresponding slot in payoff vector (what they keep)
    agentDF$payoff[as.numeric(agent)] <- agentDF$payoff[as.numeric(agent)] + (maxContribution - contribution)
    
  }
  
  contributionTot <- contributionTot * multiplier
  
  #divide the pooled contributions by the number of agents in the group
  basePay <- contributionTot/nrow(gameMatrix)
  
  agentDF$payoff[as.numeric(gamers)] <- agentDF$payoff[as.numeric(gamers)] + basePay
  
  #return the payoff vector
  return(agentDF)
  
}



#Reproduce#

#trying out different option for reproduction (sexual rep)

# SEXUAL REPRODUCTION #
reproduce_sex <- function(agentDF, selStrength, n){
  
  #rescale payoffs to be between 0-1
  agentDF$payoff <- agentDF$payoff - min(agentDF$payoff)
  agentDF$payoff <- agentDF$payoff/max(agentDF$payoff)
  
  #add constant to prob weight to fix strength of selection
  agentDF$payoff <- agentDF$payoff + (1/selStrength) #the bigger the decimal, the larger the diff in probability
  
  #make empty offspring dataframe
  offspringDF <- agentDF
  offspringDF[] <- NA
  
  for(i in 1:nrow(offspringDF)){
    
    #sample 2 agents to make love with prob based on payoff
    parents <- agentDF[sample(1:nrow(agentDF), 2, replace = F, prob = agentDF$payoff),]
    
    #inherit each weight trait from either parent
    offspringDF$A[i] <- sample(parents$A,1,replace = T)
    offspringDF$G[i] <- sample(parents$G,1,replace = T)
    offspringDF$P[i] <- sample(parents$P,1,replace = T)
    offspringDF$D[i] <- sample(parents$D,1,replace = T)
    
    #idk about association value, just gonna do the same for now
    offspringDF$association[i] <- sample(parents$association,1,replace = T)
    
  }
  
  
  #Reset payoff vec
  offspringDF$payoff <- 0
  
  
  #Mutation:
  
  #for each agent...
  for (a in 1:nrow(offspringDF)) {
    
    #change their association value to be within certain SD of original parent AV
    association_value <- rnorm(1, mean = offspringDF$association[a], sd = .05)
    
    #don't let association value get above 1
    association_value <- ifelse(association_value > 1, 1 - (association_value - 1), association_value)
    
    #don't let association value get below 0
    association_value <- ifelse(association_value < 0, 0 + abs(association_value), association_value)
    
    #assign the new mutated association value
    offspringDF$association[a] <- association_value
    
    
    #mutate prims
    #for each of the four strategy columns...
    for (c in 2:5) {
      
      #see whether or not to flip the value based on a certain prob
      random_flip <- rbinom(1, 1, flipProb)
      
      #if random flip is true, then flip the value
      offspringDF[a, c] <- ifelse(random_flip == 1, 1 - offspringDF[a, c], offspringDF[a, c])
      
    }
    
  }
  
  return(offspringDF)
  
}


# ASEXUAL REPRODUCTION #
#agents reproduce using roulette wheel selection, higher payoff = better reproductive success

reproduce_asex <- function(agentDF, selStrength, n){
  
  #rescale payoffs to be between 0-1
  agentDF$payoff <- agentDF$payoff - min(agentDF$payoff)
  agentDF$payoff <- agentDF$payoff/max(agentDF$payoff)
  
  #add constant to prob weight to fix strength of selection
  agentDF$payoff <- agentDF$payoff + (1/selStrength) #the bigger the decimal, the larger the diff in probability
  
  #generate random offspring
  offspringDF <- agentDF[sample(1:nrow(agentDF), n, replace = T, prob = agentDF$payoff),] #the bigger the decimal, the larger the diff
  
  #Reset IDs
  row.names(offspringDF) <- 1:n
  
  #Reset payoff vec
  offspringDF$payoff <- 0
  
  #Mutation:
  
  #for each agent...
  for (a in 1:nrow(offspringDF)) {
    
    #change their association value to be within certain SD of original parent AV
    association_value <- rnorm(1, mean = offspringDF$association[a], sd = .05)
    
    #don't let association value get above 1
    association_value <- ifelse(association_value > 1, 1 - (association_value - 1), association_value)
    
    #don't let association value get below 0
    association_value <- ifelse(association_value < 0, 0 + abs(association_value), association_value)
    
    #assign the new mutated association value
    offspringDF$association[a] <- association_value
    
    #mutate the weight of each primitive
    offspringDF$A[a] <- rnorm(1, mean = offspringDF$A[a], sd = .05)
    offspringDF$G[a] <- rnorm(1, mean = offspringDF$G[a], sd = .05)
    offspringDF$P[a] <- rnorm(1, mean = offspringDF$P[a], sd = .05)
    offspringDF$D[a] <- rnorm(1, mean = offspringDF$D[a], sd = .05)
    
    #don't let the weight be negative
    offspringDF$A[a] <- ifelse(offspringDF$A[a] < 0, 0, offspringDF$A[a])
    offspringDF$G[a] <- ifelse(offspringDF$G[a] < 0, 0, offspringDF$G[a])
    offspringDF$P[a] <- ifelse(offspringDF$P[a] < 0, 0, offspringDF$P[a])
    offspringDF$D[a] <- ifelse(offspringDF$D[a] < 0, 0, offspringDF$D[a])
    
    }
  
  return(offspringDF)
  
}


```



### Life Cycle

```{r lifeCycle}

start.time <- Sys.time()


#make the starting agent DFs
agentDFs <- lapply(1:modelLoops, function(x) agentGenerate(popSize))


#set up parallel processing
numCores <- detectCores()
cl <- makeCluster(numCores - 2)
registerDoParallel(cl)


#Start of life cycle#

results <- foreach(m = 1:length(agentDFs), .combine = rbind) %dopar% {
  
  resultsDF <- data.frame("loop" = rep(m),
                      "generation" = rep(0:generations),
                      "A_mean" = NA,
                      "G_mean" = NA,
                      "P_mean" = NA,
                      "D_mean" = NA
                      )
  
  
  #Generate parent agents
  agents <- agentDFs[[m]]
  
  g <- 1
  
  while(g <= generations){
    
    #make association matrix which gives agents slightly different perceptions of each others' AVs
    associationMat <- makeAssociationMat(agents, AV_SD, popSize)
  
    #generate adjacency matrix (populate with 1)
    pursuitMatrix <- matrix(1, nrow = popSize, ncol = popSize)
    
    #store each reassessment period's cumulative adjacency matrix so that agents can reference the previous one when calculating their overlap/reciprocal scores
    pursuitMatrix_list <- list(pursuitMatrix)
    
    #initialize list to store each reassessment's adjacency matrix as binary for that iteration so we can look at who was reciprocal at each reassessment period
    tiesMatrix_list <- list(length = reassessments)
  
    
    #for each reassessment... (agents look backwards so start from 2)
    for (t in 2:reassessments) {
      
      #create the empty noncumulative adjacency matrix
      tiesMatrix <- matrix(0, nrow = popSize, ncol = popSize)
    
      #Calculate overlap in previous adjacency matrix#
      overlap_matrices <- overlap_prims(pursuitMatrix_list[[t-1]])
      
      #Each agent sends out pursuits#
      
      #for each agent...
      for (a in 1:nrow(agents)) {
        
        #calculate overlap scores between themselves and the other agents
        overlap_scores <- overlap_mean(a, agents, overlap_matrices)
        
        #calculate the reciprocity scores between themselves and the other agents
        reciprocity_scores <- reciprocity(a, pursuitMatrix_list[[t - 1]])
        
        #calculate the association values
        association_values <- associationMat[a,]
        
        #Combine overlap, reciprocity, and association scores into single calculation
        pursuit_scores <-overlap_scores * reciprocity_scores * association_values
        
        #sort by product in descending order and get the indices of the top value agents
        pursuit_indices <- order(-pursuit_scores)[1:pursuitAmount]
        
        #add 1 to their pursuits in the cumulative the adjacency matrix
        pursuitMatrix[a, pursuit_indices] <- pursuitMatrix[a, pursuit_indices] + 1
        
        #and to the individual one
        tiesMatrix[a, pursuit_indices] <- 1
        
        
      }
      
      #add the pursuitMatrix to the matrix list
      pursuitMatrix_list[[t]] <- pursuitMatrix
      
      #add the non cumulative matrix to its list
      tiesMatrix_list[[t]] <- tiesMatrix
      
      #make matrix that has only reciprocated ties
      friendsMatrix <- tiesMatrix * t(tiesMatrix)
      
      
      #Game#
      
      #create empty payoff vector to store payoffs (later used for reproduction)
      #payoffs <- vector(mode = "numeric", length = popSize)
      
        #randomly pick one agent to play game
        gamer <- sample(popSize, 1)
      
        for (i in sample(1:popSize)) {
        #randomly pick one agent to play game
        gamer <- i
        
        #pull indices of agents' reciprocated ties
        gamerGroup <- which(friendsMatrix[gamer, ] == 1)
        
        #add the gamer to the group indices
        gamerGroup <- c(gamerGroup, gamer)
        
        #subset the ties matrix to include only those in group
        gamersMat <- friendsMatrix[gamerGroup, gamerGroup, drop = F]
        
        #change the names back to original indices
        rownames(gamersMat) <- gamerGroup
        colnames(gamersMat) <- gamerGroup
        
        #Agents decide whether to play a PG game proportional to number of reciprocated ties in group
        #Game happens only if all agents agree
        if (nrow(gamersMat) == (pursuitAmount + 1) &
            prod(rbinom(nrow(gamersMat), 1, rowSums(gamersMat) / nrow(gamersMat))) == 1) {
          #calculate the payoff each agent will receive
          agents <- payoff(gamersMat, gamerGroup, agents)
          
        }
        
      }
      
    }
    

    #Reproduce#
    
    #agents reproduce
    agents <- reproduce_asex(agents, selStrength, popSize) 
    
    #get the average weight of each prim across agents
    avWeight_A <- mean(agents$A)
    avWeight_G <- mean(agents$G)
    avWeight_P <- mean(agents$P)
    avWeight_D <- mean(agents$D)
    
    #add these to resultsDF DF
    resultsDF$A_mean[resultsDF$generation == g] <- avWeight_A
    resultsDF$G_mean[resultsDF$generation == g] <- avWeight_G
    resultsDF$P_mean[resultsDF$generation == g] <- avWeight_P
    resultsDF$D_mean[resultsDF$generation == g] <- avWeight_D
    
    g <- g+1
    
    print(g)
    
  }
  
  resultsDF
  
}
  
#end parallel processing
stopCluster(cl)



end.time <- Sys.time()

print(end.time - start.time)

#write.csv(results, myPath)
      
```

### Analysis

```{r analysis}


#put into into long format
results_long <- pivot_longer(results, cols = c(3:6), names_to = "prim", values_to = "mean_true")

#plot 
p1 <- ggplot(data = results_long,
               aes(x = generation,
                   y = mean_true,
                   color = prim)) +
  labs(x = "Generation",
       y = "Mean agents using prim",
       color = "Phenotype") +
  geom_line() +
#  xlim(min(results$generation), max(results$generation) + 0.5) + 
  theme_minimal() +
  facet_wrap(~loop)

plot(p1)


###

#adding prim labels directly to the line

# results_labs <- results %>%
#   group_by(Var1) %>%
#   filter(generation == max(generation)) %>%
#   ungroup()

  # geom_text(
  #   data = results_labs,
  #   aes(label = prim),
  #   hjust = -0.1,
  #   size = 4,
  #   fontface = "bold"
  #) +

```
















