---
title: "groupFormation"
author: "Clara Lavrador"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Overview

### Purpose


### Agents

Each agent will have a set of three traits and three preferences, which are randomly assigned from 0-100. 
They will also have a number of pursuits they can make per regrouping period.


### Life Cycle

1. Gather info
2. Pursue agents
3. Regroup

#### Gather info

Each agent will calculate three scores with respect to other agents: association value, reciprocity score, and overlap score.

The association value is calculated by taking the euclidean distance of the agent's preferences with the other agents' traits. 

The reciprocity score is the proportion of pursuits between the agent and each other agent that are reciprocated.

The overlap score is a measurement of pursuit similarity (both sent and received) between the agent and each other agent, with respect to third parties.
  It is based on David Pietraszewski's triadic primitives, whereby:
   
     alliance is the correlation between the agent and each other agents' sent pursuits (i.e., how much they pursue the same agents)
     
     generalization is the correlation between the agent and each other agents' received pursuits (i.e., how much they are pursued by the same agents)
     
     displacement is the correlation between the agent's sent pursuits and each other agents' received pursuits (i.e., how much the agents that agent b pursues pursue agent a)
     
     defense is the correlation between the agent's received and each other agents' sent pursuits (i.e., how much the agents that agent a is pursued by are pursued by agent b)

  
These three scores are then formulated to produce a single pursuit score:

  pursuit value = association value + (reciprocity score + overlap score)


#### Pursue agents

Each agent will pursue the agents with the highest pursuit values. 

Sent and received pursuits will be accumulated (and thus used to calculate overlap in the next reassessment period) 

#### Regroup

Agents who have reciprocated pursuits with each other will be considered "grouped" together for that reassessment period.

Then it all begins anew until the max number of reassessment periods has been reached.


### Analysis

For now I'm plotting the number of reciprocated edges over time across all model loops.
But I must think of a better representation... number of groups only tells half the story (bc they still differ in size)

Best thing would be to see if the model loops converge in some other aspect(s)

Also I could compare it to a model that does not include overlap score



## Model

### Packages 
```{r packages}

library(ggplot2)
library(igraph)
library(network)
library(networkDynamic)
library(intergraph)
library(ndtv)
library(dplyr)
library(doParallel)

```


### Parameters
```{r parameters}

#number of agents to generate
popSize <- 100

#number of pursuits each agent will have for each reassessment period
pursuitAmount <- 3

#number of times agents will reevaluate their pursuits
reassessments <- 200

#number of loops for the whole model
modelLoops <- 1

#number of generations
generations <- 1

#SD of association value of agents in the eyes of other agents
AV_SD <- 0.05

#set strength of selection
selStrength <- 15

#multiplier
multiplier <- 1.2

#mutation probability that binary prim trait will turn on/off 
flipProb <- 0.01

#percent of pursuits that will not be visible to agent
obscurePercent <- 0.75

myPath <- "C:/Users/clavra/Documents/groupFormation/resultsMain.csv"

```


### Functions
```{r functions}

#Agent Generate Function#

agentGenerate <- function(n){
  
  #generate IDs for each agent
  #ID <- 1:n
  
  #generate association value
  association <- abs(rnorm(n, mean = 50, sd = 10))
  association <- round(association, digits = 0)
  association <- (association/100)
  
  #generate weight of each prim for each agent 
  A <- rbinom(n, 1, 0.5)
  
  G <- rbinom(n, 1, 0.5)
  
  P <- rbinom(n, 1, 0.5)
  
  D <- rbinom(n, 1, 0.5)
  
  payoff <- 0
  
  #put it all together in a df
  agents <- data.frame(association, A, G, P, D, payoff)
    
  
  return(agents)
  
}


#Association Matrix#
#each agent will see each other agent's association value slightly differently

makeAssociationMat <- function(agentDF, sd, n){
  
  associationMat <- matrix(, nrow = n, ncol = n)
  
  for(a in 1:nrow(agentDF)){
    
    associationMat[,a] <-  rnorm(n, mean = agentDF$association[a], sd = sd)
    
  }
  
  diag(associationMat) <- NA
  
  return(associationMat)
  
}
  
  
#Weighted Jaccard#
#used to calculate overlap#

weighted_jaccard <- function(x, y) {
  
  # Calculate the numerator: Sum of minimum intensities
  numerator <- sum(pmin(x, y))
  
  # Calculate the denominator: Sum of maximum intensities
  denominator <- sum(pmax(x, y))
  
  # Handle edge case where denominator is 0 (all zeros)
  if (denominator == 0) return(0)
  
  # Calculate weighted Jaccard similarity
  similarity <- numerator / denominator
  return(similarity)
}


#Overlap Score Calculation

overlap <- function(agent, agentDF, adjMat){

  #initialize component df
    scores <- list(
      alliance = NA,
      displacement = NA,
      defense = NA,
      generalization = NA
    )
    
    if(sum(agentDF[agent,c("A","G","P","D")]) == 0)
      return(rep(NA, popSize))
    
    #Alliance# (row-row)
    
    #if the agent's strat includes alliance...
    if(agentDF$A[agent] == 1){
      
      #calculate weighted jaccard with each row against against agent's row
      scores$alliance <- sapply(1:nrow(adjMat), function(i) {
        
        #skip own row  
        if (i == agent)
          return(NA)
          
        #ignore the cells in the row of the agent and current target agent
        exclude_indices <- c(agent, i)
        
        #get target's row not including own and agent cell
        x_filtered <- adjMat[i,-exclude_indices]
        
        #get agent's row not including own and target cell
        agent_filtered <- adjMat[agent,-exclude_indices]
        
        #calculate weighted jaccard on those rows
        weighted_jaccard(x_filtered, agent_filtered)
        
      })
    }
      
    
    #Generalization# (col-col)
      
    #if the agent's strat includes generalization
    if(agentDF$G[agent] == 1){
      
      #calculate weighted jaccard with each col against against agent's col
      scores$generalization <- sapply(1:ncol(adjMat), function(j) {
        
        #skip own col  
        if (j == agent)
          return(NA)
          
        #ignore the cells in the col of the agent and current target agent
        exclude_indices <- c(agent, j)
        
        #get target's col not including own and agent cell
        x_filtered <- adjMat[-exclude_indices,j]
        
        #get agent's col not including own and target cell
        agent_filtered <- adjMat[-exclude_indices,agent]
        
        #calculate weighted jaccard on those cols
        weighted_jaccard(x_filtered, agent_filtered)
      })
    }
      
    #Displacement# (row-col)
      
    #if the agent's strat includes displacement
    if(agentDF$P[agent] == 1){
      
      #calculate weighted jaccard with each col against agent's row
      scores$displacement <- sapply(1:ncol(adjMat), function(j) {
        
        #skip own col  
        if (j == agent)
          return(NA)
          
        #ignore the cells in the col of the agent and current target agent
        exclude_indices <- c(agent, j)
        
        #get target's col not including own and agent cell
        x_filtered <- adjMat[-exclude_indices,j]
        
        #get agent's row not including own and target cell
        agent_filtered <- adjMat[agent, -exclude_indices]
        
        #calculate weighted jaccard on row and col
        weighted_jaccard(x_filtered, agent_filtered)
      })
    }
      
    #Defense# (col-row)
      
    #if the agent's strat includes defense
    if(agentDF$D[agent] == 1){
      
      #calculate weighted jaccard with each row against agent's col
      scores$defense <- sapply(1:nrow(adjMat), function(i) {
        
        #skip agent's row  
        if (i == agent)
          return(NA)
          
        #ignore the cells of the agent and current target agent
        exclude_indices <- c(agent, i)
        
        #get target's row not including own and agent cell
        x_filtered <- adjMat[i,-exclude_indices]
        
        #get agent's col not including own and target cell
        agent_filtered <- adjMat[-exclude_indices,agent]
        
        #calculate weighted jaccard between agent's col and others' row
        weighted_jaccard(x_filtered, agent_filtered)
        
      })
    }
    
    selected_scores <- do.call(rbind, scores)
    
    #if we doing mean
    #overlap_scores <- colMeans(selected_scores, na.rm = TRUE)
    
    #if we doing min
    overlap_scores <- apply(selected_scores, 2, FUN = min)
    
    overlap_scores[agent] <- NA
    
    return(overlap_scores)
  
  
}


#Reciprocity Function#

reciprocity <- function(agent, adjMat){
  
  #pull agent's column which represents their received pursuits
  received <- adjMat[,agent]
  
  #pull agent's row which represents their sent pursuits
  sent <- adjMat[agent,]
    
  #calculate proportions of pursuits which were received
  reciprocity_scores <- (received / (received + sent))
  
  #nullify their own reciprocity score
  reciprocity_scores[agent] <- NA
  
  #turn NaNs into 0 
  reciprocity_scores[is.nan(reciprocity_scores)] <- 0
  
  return(reciprocity_scores)
  
}


#Contributions#

payoff <- function(gameMatrix, gamers, agentDF){
  
  #start the sum
  contributionTot <- 0
  
  #for each agent...
  for(agent in gamers){
    
    #their max contribution is their association value
    maxContribution <- agentDF$association[row.names(agentDF) == agent] * 100
    
    #get the percent of the AV they will contribute, which is proportional to number of recip ties within group
    proportionTied <- if (nrow(gameMatrix) > 1) sum(gameMatrix[as.character(agent), ])/(nrow(gameMatrix)-1) else 0
    
    
    #calculate their contribution 
    contribution <- maxContribution * proportionTied
    
    #add their total contribution to those of the other agents
    contributionTot <- contributionTot + contribution
    
    #add their noncontributed pay to their corresponding slot in payoff vector (what they keep)
    agentDF$payoff[as.numeric(agent)] <- agentDF$payoff[as.numeric(agent)] + (maxContribution - contribution)
    
  }
  
  contributionTot <- contributionTot * multiplier
  
  #divide the pooled contributions by the number of agents in the group
  basePay <- contributionTot/nrow(gameMatrix)
  
  agentDF$payoff[as.numeric(gamers)] <- agentDF$payoff[as.numeric(gamers)] + basePay
  
  #return the payoff vector
  return(agentDF)
  
}




#Reproduce#

# ASEXUAL REPRODUCTION #
#agents reproduce using roulette wheel selection, higher payoff = better reproductive success
reproduce <- function(agentDF, selStrength, n){
  
  #rescale payoffs to be between 0-1
  agentDF$payoff <- agentDF$payoff - min(agentDF$payoff)
  agentDF$payoff <- agentDF$payoff/max(agentDF$payoff)
  
  #add constant to prob weight to fix strength of selection
  agentDF$payoff <- agentDF$payoff + (1/selStrength) #the bigger the decimal, the larger the diff in probability
  
  #generate random offspring
  offspringDF <- agentDF[sample(1:nrow(agentDF), n, replace = T, prob = agentDF$payoff),] #the bigger the decimal, the larger the diff
  
  #Reset IDs
  row.names(offspringDF) <- 1:n
  
  #Reset payoff vec
  offspringDF$payoff <- 0
  
  #Mutation:
  
  #for each agent...
  for (a in 1:nrow(offspringDF)) {
    
    #change their association value to be within certain SD of original parent AV
    association_value <- rnorm(n, mean = offspringDF$association[a], sd = .03)
    
    #don't let association value get above 1
    association_value <- ifelse(association_value > 1, 1 - (association_value - 1), association_value)
    
    #don't let association value get below 0
    association_value <- ifelse(association_value < 0, 0 + abs(association_value), association_value)
    
    #assign the new mutated association value
    offspringDF$association[a] <- association_value
    
    #flip each trait or nah and replace into corresponding cells
    offspringDF[a, 2:5] <- (offspringDF[a, 2:5] + rbinom(4, 1, flipProb)) %% 2
    
   }
  
  return(offspringDF)
  
}

#shift fiveMatsList to discard oldest mat and add newest
append_shift <- function(lst, newMat) {
  
  #drop oldest mat, append newest
  lst <- c(lst[-1], list(newMat))  
    
  return(lst)
    
}


#get the flat indices of the agent's rows and cols in matrix (used so we don't obscure direct ties)
getAgentIxs <- function(agent, matrix){
  
  #get flat inds of agent's col
  startC <- (agent-1)*nrow(matrix) + 1
  endC <- startC + (nrow(matrix)-1)
  
  colIndices <- startC:endC
  
  #get flat inds of agent's row
  rowIndices <- seq(from = agent, by = nrow(matrix), length.out = nrow(matrix))
  
  #bind together the row and col inds
  allIxs <- c(colIndices, rowIndices)
  
  return(allIxs)
  
}


#Making Noisy Ties Mat#
makeNoisyNewTies <- function(agent, godTiesMatrix, obscurePercent) {
    
    #make agent's ties matrix based on overall ties mat for this round
    noisyNewTies <- godTiesMatrix
    
    #get the indicies of the agent's direct pursuits for us to ignore when noisifying
    agentIxs <- getAgentIxs(agent, godTiesMatrix)
    
    #get all indices where there is a 1
    positivePursuits <- which(godTiesMatrix == 1, arr.ind = F)
    
    #remove the 1 indices that are the agent's direct pursuits
    indPositivePursuits <- positivePursuits[!positivePursuits %in% agentIxs]
    
    #get the indices of the 75% of indirect 1s that we are gonna turn into 0s
    obscuredPursuits <- sample(indPositivePursuits, obscurePercent * length(indPositivePursuits), replace = F)
    
    #turn those 1s into 0s
    noisyNewTies[obscuredPursuits] <- 0
    
    return(noisyNewTies)
    
}







```



### Life Cycle

```{r lifeCycle}

start.time <- Sys.time()

results <- data.frame("loop" = rep(1:modelLoops, each = generations+1),
                      "generation" = rep(0:generations),
                      "A_mean" = NA,
                      "G_mean" = NA,
                      "P_mean" = NA,
                      "D_mean" = NA
                      )




for(m in 1:modelLoops) {
  
  print(m)
  
  #Generate parent agents
  agents <- agentGenerate(popSize)
  
  #add the initial prim means to generation 0 in results
  results[results$loop == m & results$generation == 0, c(3:6)] <- colMeans(agents[2:5])
  
  
  #set first generation
  g <- 1
  
  while(g <= generations){
    
    #make association matrix which gives agents slightly different perceptions of each others' AVs
    associationMat <- makeAssociationMat(agents, AV_SD, popSize)
    
    #initiate the 5 individual ties mats and store them in a list of lists
    fiveMatsLists <- replicate(popSize, simplify = F, 
                                    replicate(5, matrix(0, nrow = popSize, ncol = popSize), 
                                              simplify = FALSE))
  
    #initiate the individual pursuit mats
    indivPursuits <- replicate(popSize, simplify = F, matrix(0, nrow = popSize, ncol = popSize))
    
    
    #for each reassessment... (agents look backwards so start from 2)
    for (t in 2:reassessments) {
      
    #create matrix that holds every definitive pursuit for this round
    godTiesMatrix <- matrix(0, nrow = popSize, ncol = popSize)
      
      #Each agent sends out pursuits#
      for (a in 1:nrow(agents)) {
        
        #calculate overlap scores between themselves and the other agents
        overlap_scores <- overlap(a, agents, indivPursuits[[a]])
        
        #calculate the reciprocity scores between themselves and the other agents
        reciprocity_scores <- reciprocity(a, indivPursuits[[a]])
        
        #calculate the association values
        association_values <- associationMat[a,]
        
        #Combine overlap, reciprocity, and association scores into single calculation
        pursuit_scores <- colMeans(rbind(overlap_scores, reciprocity_scores, association_values), na.rm = T)
        
        #sort by product in descending order and get the indices of the top value agents
        pursuit_indices <- order(-pursuit_scores)[1:pursuitAmount]
        
        #update the overall ties matrix which shows each agent's pursuit for this round
        godTiesMatrix[a, pursuit_indices] <- 1
    
      }
      
      #make matrix that has only reciprocated ties
      friendsMatrix <- godTiesMatrix * t(godTiesMatrix)
      
    
      #Updating ind pursuit mat and five mats to be used in next round
      
      for(a in 1:popSize){
        
        #Update indivPursuitMat
        
        #make agents forget the info from last 5th round
        indivPursuits[[a]] <- indivPursuits[[a]] - fiveMatsLists[[a]][[1]]
        
        #create new noisy ties from latest round
        newTies <- makeNoisyNewTies(a, godTiesMatrix, obscurePercent)
        
        #add new ties info to their indivPursuitMat
        indivPursuits[[a]] <- indivPursuits[[a]] + newTies
        
        
        #Update fiveMatsList
        
        #shift the 5 list to discard oldest and add newTies
        fiveMatsLists[[a]] <- append_shift(fiveMatsLists[[a]], newTies)
        
      }
      
      
      #Game#
      
      #create empty payoff vector to store payoffs (later used for reproduction)
      #payoffs <- vector(mode = "numeric", length = popSize)
      
        #randomly pick one agent to play game
        gamer <- sample(popSize, 1)
      
        for (i in sample(1:popSize)) {
        #randomly pick one agent to play game
        gamer <- i
        
        #pull indices of agents' reciprocated ties
        gamerGroup <- which(friendsMatrix[gamer, ] == 1)
        
        #add the gamer to the group indices
        gamerGroup <- c(gamerGroup, gamer)
        
        #subset the ties matrix to include only those in group
        gamersMat <- friendsMatrix[gamerGroup, gamerGroup, drop = F]
        
        #change the names back to original indices
        rownames(gamersMat) <- gamerGroup
        colnames(gamersMat) <- gamerGroup
        
        #Agents decide whether to play a PG game proportional to number of reciprocated ties in group
        #Game happens only if all agents agree
        if (nrow(gamersMat) == (pursuitAmount + 1) &
            prod(rbinom(nrow(gamersMat), 1, rowSums(gamersMat) / nrow(gamersMat))) == 1) {
          
          #calculate the payoff each agent will receive
          agents <- payoff(gamersMat, gamerGroup, agents)
          
        }
        
      }
      
    }
    

    #Reproduce#
    
    #save parents
    parents <- agents
    
    #agents reproduce
    agents <- reproduce(agents, selStrength, popSize) 
    
    #get the average weight of each prim across agents
    avWeight_A <- mean(agents$A)
    avWeight_G <- mean(agents$G)
    avWeight_P <- mean(agents$P)
    avWeight_D <- mean(agents$D)
    
    #add these to results DF
    results$A_mean[results$generation == g & results$loop == m] <- avWeight_A
    results$G_mean[results$generation == g & results$loop == m] <- avWeight_G
    results$P_mean[results$generation == g & results$loop == m] <- avWeight_P
    results$D_mean[results$generation == g & results$loop == m] <- avWeight_D

    print(results[results$generation == g & results$loop == m, 3:6])
    
    g <- g+1
    
    print(g)
    
  }
  
  end.time <- Sys.time()
}
  

#remember that ties mat has all pursuits, not just reciprocal



end.time <- Sys.time()

print(end.time - start.time)

#write.csv(results, myPath)
      
```

### Analysis

```{r analysis}

#put into into long format
results_long <- pivot_longer(results, cols = c(3:6), names_to = "prim", values_to = "mean_true")


p1 <- ggplot(data = results_long,
               aes(x = generation,
                   y = mean_true,
                   color = prim)) +
  labs(x = "Generation",
       y = "Mean agents using prim",
       color = "Phenotype") +
  geom_line() +
  # geom_text(
  #   data = results_labs,
  #   aes(label = prim),
  #   hjust = -0.1,
  #   size = 4,
  #   fontface = "bold"
  #) +
#  xlim(min(results$generation), max(results$generation) + 0.5) +  # Give space for labels
  theme_minimal() +
  facet_wrap(~loop)



plot(p1)


#####

lastGraph <- graph_from_adjacency_matrix(friendsMatrix, mode = "undirected")

plot(lastGraph, 
     layout = layout.fruchterman.reingold,
   #  vertex.label = NA,
     vertex.size = 9
     )


```

















